---
title: "Advanced Business Analytics using R Project"
author: "Mihir Raikar"
date: "4/16/2019"
output: pdf_document
---

```{r Loading Libraries}
library(data.table)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(e1071)
library(caret)
library(tidyverse)
library(gains)
library(corrplot)
library(mefa4)
library(randomForest)
library(glmnet)
library(corrplot)
library(gridExtra)
library(arm)
library(GGally)
library(car)
library(ggcorrplot)
```



```{r Loading the Data}
wine <- fread("winequality-white.csv")
```



```{r Data at a Glance}
sum(duplicated(wine))
dim(wine)
summary(wine)
```


Our Dependent Variable is Quality, hence we will take a look at the distribution of quality
```{r Plot Quality}
quality_plot <- ggplot(aes(quality), data = wine) + geom_bar() + ggtitle ("Quality chart")
quality_plot
```


The range of Quality is from 3 to 9. We can bucket the data into good, medium and bad with reference to Quality
```{r Quality Level}
wine$quality_level <- ifelse(wine$quality >= 7, "good", NA) 
wine$quality_level <- ifelse(wine$quality <= 6, "medium", wine$quality_level)
wine$quality_level <- ifelse(wine$quality <= 4, "bad", wine$quality_level)
wine$quality_level <- as.factor(wine$quality_level)
wine$quality <- as.factor(wine$quality)
```

Plot and Visualize Quality level
```{r}
qualitylevel_plot <- ggplot(aes(quality,fill=quality_level),data=wine) + geom_bar() + ggtitle ("Quality level chart")
qualitylevel_plot
qualitylevelcount_plot<-qplot(wine$quality_level) + xlab("quality level") + ggtitle("count of quality level")
grid.arrange(quality_plot,qualitylevel_plot,ncol=2)
```

We will now explore various the variables


```{r Fixed Acidity}
summary(wine$fixed_acidity)
ggplot(aes(fixed_acidity), data = wine) + geom_bar()
ggplot(aes(x=quality, y=fixed_acidity), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=fixed_acidity), data = wine) + geom_boxplot()
```
The data is positively skewed and also has some outliers
The average fixed acidity tends to decrease with increase in quality but increases with quality=9
The observations of fixed acidity at it has in general a negative relationship with quality level


```{r Volatile Acidity}
summary(wine$volatile_acidity)
ggplot(aes(volatile_acidity), data = wine) + geom_bar()
ggplot(aes(x=quality, y=volatile_acidity), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=volatile_acidity), data = wine) + geom_boxplot()
```
Similar to fixed acidity, volatile acidity is also positive skewed with outliers but the range is very small
The relationship with quality is unclear as there is no trend
The relationship of volatile acidity with quality level is negative


```{r Citric Acid}
summary(wine$citric_acid)
ggplot(aes(citric_acid), data = wine) + geom_bar()
ggplot(aes(x=quality, y=citric_acid), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=citric_acid), data = wine) + geom_boxplot()
```
The distribution of citric acid is similar to normal but has 2 unusual peaks and a few outliers.
The peaks are causing the data to not have a general trend

```{r Residual Sugar}
summary(wine$residual_sugar)
ggplot(aes(residual_sugar), data = wine) + geom_bar()
ggplot(aes(x=quality, y=residual_sugar), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=residual_sugar), data = wine) + geom_boxplot()
```
The data is positively skewed. It looks like most of the wines have very low residual sugars
It looks like residual sugar is low in bad wines but comaritively high in medium and again comparatively lower in good.
So generally, residual sugar in wine is good, but really good wines dont have it as much as average wines

```{r Chlorides}
summary(wine$chlorides)
ggplot(aes(chlorides), data = wine) + geom_bar()
ggplot(aes(x=quality, y=chlorides), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=chlorides), data = wine) + geom_boxplot()
```
Chlorides are also positively skewed. The max value is lot higher than mean.
Chlorides also have a negative relationship with quality level but not very significant

```{r Free Sulphur Dioxide}
summary(wine$free_sulfur_dioxide)
ggplot(aes(free_sulfur_dioxide), data = wine) + geom_bar()
ggplot(aes(x=quality, y=free_sulfur_dioxide), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=free_sulfur_dioxide), data = wine) + geom_boxplot()
```
There is a huge outlier(289) whereas the rest of the data is below 150. 
Apart from the outlier also, the data is a little positive skewed.
There is a positive relationship with quality but is not that significant with good quality

```{r Total Sulphur Dioxide}
summary(wine$total_sulfur_dioxide)
ggplot(aes(total_sulfur_dioxide), data = wine) + geom_bar()
ggplot(aes(x=quality, y=total_sulfur_dioxide), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=total_sulfur_dioxide), data = wine) + geom_boxplot()
```
There are some peaks but the distribution is similar to normal
Similar to residual sugar, the bad wines have low total sulphur dioxide, medium has high but good have lower than medium

```{r Density}
summary(wine$density)
ggplot(aes(density), data = wine) + geom_bar()
ggplot(aes(x=quality, y=density), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=density), data = wine) + geom_boxplot()
```
The range is very low and has a couple of outliers
There is a negative relationship between density and quality level

```{r pH}
summary(wine$volatile_acidity)
ggplot(aes(pH), data = wine) + geom_bar()
ggplot(aes(x=quality, y=pH), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=pH), data = wine) + geom_boxplot()
```
pH has a distribution very close to normal. there are peaks but no significant outliers
There is a clear positive relationship between pH and Quality/Quality level.

```{r Sulphates}
summary(wine$sulphates)
ggplot(aes(sulphates), data = wine) + geom_bar()
ggplot(aes(x=quality, y=sulphates), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=sulphates), data = wine) + geom_boxplot()
```
The data is positive skewed and a few high values
Clear positive relation with quality

```{r Alchohol}
summary(wine$alcohol)
ggplot(aes(alcohol), data = wine) + geom_bar()
ggplot(aes(x=quality, y=alcohol), data = wine) + geom_boxplot()
ggplot(aes(x=quality_level, y=alcohol), data = wine) + geom_boxplot()
```
There are a number of peaks, looks like no outliers
There seems to be a positive relationship with quality butthe trend is not completely clear. 
Good wines have very high alchohol but medium wines have lower alchohol than bad wines which is surprising

```{r Correltions}
ggcorr(wine)
```
Density and Residual sugar has really high correlation (0.84)
Density and alchohol also has a high (negative) correlation (-0.78)
Density also as a high correlation with total sulfur dioxide(0.53)
Free sulfur dioxide and total sulfur dioxide are also highly correlated(0.62)
Alchohol is negatively correlated with total sulfur dioxide and residual sugar(both -45)
Alchohol and quality has a correlation of 44
pH and fixed acidity are also negatively correlated (-0.43)


```{r Data Partitioning}
set.seed(123)
samp <- sample(nrow(wine), 0.6 * nrow(wine))
train <- wine[samp, ]
test <- wine[-samp, ]
```

```{r SVM on Quality}
svmlinear1 <- train(quality ~., data = train, 
                   method = "svmLinear",
                   trControl=trainControl(method = "repeatedcv", 
                                          number = 10, repeats = 10),  
                   preProcess = c("center", "scale"),
                   tuneLength = 10)

svmlinear1$bestTune
summary(svmlinear1)

predsvmLinear1 <- predict(svmlinear1, test)
confusionMatrix(table(predsvmLinear1, test$quality))
```
We have run the SVM model on Quality variable. For the best model, we found C=1. The accuracy is 74.29%

```{r SVM on Quality Level}
svmlinear2 <- train(quality_level ~., data = train, 
              method = "svmLinear",
              trControl=trainControl(method = "repeatedcv", 
                                     number = 10, repeats = 10),  
              preProcess = c("center", "scale"),
              tuneLength = 10)

svmlinear2$bestTune
summary(svmlinear2)

predsvmLinear2 <- predict(svmlinear2, test)
confusionMatrix(table(predsvmLinear2, test$quality_level))
```
We have run the SVM model on Quality Level variable. For the best model, we found C=1. The accuracy is 100%

```{r Naive Bayes on Quality}
nb1 <- naiveBayes(quality ~ ., data = train)
nb1

# probabilities
pred.prob <- predict(nb1, newdata = test, type = "raw")
# class membership
pred.class1 <- predict(nb1, newdata = test)
confusionMatrix(pred.class1, test$quality)
```
We have run the Naive Bayes model on Quality variable. The accuracy is 70.2%

```{r Naive Bayes on Quality Level}
nb2 <- naiveBayes(quality_level ~ ., data = train)
nb2

# probabilities
pred.prob <- predict(nb2, newdata = test, type = "raw")
# class membership
pred.class2 <- predict(nb2, newdata = test)
confusionMatrix(pred.class2, test$quality_level)
```
We have run the Naive Bayes model on Quality Level variable. The accuracy is 97.81%

```{r Ridge on Quality}
set.seed(123)
ridgeReg1 <- train(quality~., train, method = 'glmnet',
                  tuneGrid = expand.grid(alpha = 0, 
                                         lambda = seq(0.0001, 0.5, length = 5)),
                  trControl = trainControl(method = "repeatedcv", 
                                           number = 10, repeats = 10,
                                           verboseIter = TRUE))
# print results
print(ridgeReg1)

# plot results
plot(ridgeReg1)
plot(varImp(ridgeReg1, scale = FALSE))
plot(varImp(ridgeReg1, scale = TRUE))

PredictRidge1 <- predict(ridgeReg1, test)
confusionMatrix(PredictRidge1, test$quality)
```
We have run the Ridge model on Quality variable. The accuracy is 71.99%

```{r Ridge on Quality Level}
set.seed(123)
ridgeReg2 <- train(quality_level~., train, method = 'glmnet',
                  tuneGrid = expand.grid(alpha = 0, 
                                         lambda = seq(0.0001, 0.5, length = 5)),
                  trControl = trainControl(method = "repeatedcv", 
                                           number = 10, repeats = 10,
                                           verboseIter = TRUE))
# print results
print(ridgeReg2)

# plot results
plot(ridgeReg2)
plot(ridgeReg2$finalModel, xvar = 'lambda', lwd =1.4, label = TRUE)
plot(varImp(ridgeReg2, scale = FALSE))
plot(varImp(ridgeReg2, scale = TRUE))

PredictRidge2 <- predict(ridgeReg2, test)
confusionMatrix(PredictRidge2, test$quality_level)

```
We have run the Ridge model on Quality Level variable. The accuracy is 99.69%

```{r Lasso on Quality}
set.seed(123)
lassoReg1 <- train(quality~., train, method = 'glmnet',
                  tuneGrid = expand.grid(alpha = 1, 
                                         lambda = seq(0.0001, 0.5, length = 5)),
                  trControl = trainControl(method = "repeatedcv", 
                                           number = 10, repeats = 5,
                                           verboseIter = TRUE))

# plot results
plot(lassoReg1)
plot(varImp(lassoReg1, scale = FALSE))
plot(varImp(lassoReg1, scale = TRUE))

PredictLasso1 <- predict(lassoReg1, test)
confusionMatrix(PredictLasso1, test$quality)
```
We have run the Lasso model on Quality variable. The accuracy is 73.42%

```{r Lasso on Quality Level}
lassoReg2 <- train(quality_level~., train, method = 'glmnet',
                  tuneGrid = expand.grid(alpha = 1, 
                                         lambda = seq(0.0001, 0.5, length = 5)),
                  trControl = trainControl(method = "repeatedcv", 
                                           number = 10, repeats = 5,
                                           verboseIter = TRUE))
# plot results
plot(lassoReg2)
plot(lassoReg2$finalModel, xvar = 'lambda', lwd =1.4, label=TRUE)
plot(varImp(lassoReg2, scale = FALSE))
plot(varImp(lassoReg2, scale = TRUE))

PredictLasso2 <- predict(lassoReg2, test)
confusionMatrix(PredictLasso2, test$quality_level)
```
We have run the Lasso model on Quality Level variable. The accuracy is 100%



